---
title: "Gender bias at ASM journals"
author: "Ada Hagan"
date: "July 17, 2018"
output:
  html_document:
    df_print: paged
csl: asm.csl
bibliography: references.bib
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(markdown)
library(rmarkdown)
library(formattable)

source("code/load_data.R")
knitr::opts_chunk$set(echo = TRUE)
```

#Abstract

#Introduction

  + Gender disparity in stem
  
  + Studies evaluating gender in publications
      + Nature
      + JAMA
      + Glam journals
      + Science?
      + Ecology (Cho, 2014; Pettorelli, 2013; Buckley, 2014)
      + Medical (Amrein, 2011; del Carmen, 2003)
      + Authorship analyis of 8 million papers - West, 2013
      + Reviewer practice by gender - Grod, 2008
      + recognition for group work - Sarsons, 2017
  
  + Few look at submissions to understand source of disparity
      + AGU
      + JAMA (Gilbert & Williams)
      + (Buckley, 2014)
      + elife preprint
  
  + This study
      + Cumulative data since 2012 to see where ASM stands
      + Use submissions and compare to membership to see where disparity exists (if at all)

#Results
```{r, plot_rej_rates} 
source("code/reject_by_gender.R")

auth_type_A #cross ASM rejection rates (all genders)
auth_type_B #rejection rates by journ (m/w) -- need to highlight instances with diff > 5%
auth_type_C #breakdown of submitted manuscripts (% by gender) according to EJP.decision
auth_type_D #breakdown of C by journal -- need to highlight instances with diff > 5%
auth_type_E #breakdown of small value decisions
```
  + higher rejection rates for women than men, esp corresponding authors
  + move forward examining corresponding authors only b/c greatest effect


  + Results:
      + Women recieve more editorial rejections than men, proportionally (editorial_rejections.R)
    + time_to_publication.R:
      + Papers published by women take slightly longer (from submission to ready for publication) than men at some journals (mSphere, mBio, mSystems, CVI, JB, JCM, AEM)
      + but do not require a greater number of revisions to be accepted than men
      + Papers rejected following review that were submitted by women do not take longer (in days) to be rejected or have more revisions (do they move on more quickly?) -- Why the F are people submitting 5+ versions before being rejected?? (time_to_final_rejection.R)
      + Decisions are returned to men & women similarly (time_at_journ.R)
      + Slight disparity in gendered sumissions to high (mBio) vs lower (mSphere) journals (bias_by_jif.R)
    + impact_analysis.R:
      + Women tend to receive lower cites per month published than men 
      + HTML/PDF/Abstract views are equivalent 
      + women are more likely to be published in lower JIF journals
    + rej_by_inst_type.R:
      + slight disparity in editorial rejection rates according to institution type, greatest difference (4%) occurs for institute/medical schools
      + trend holds for most journals and is >20% diff at JCM (R2s), mBio (Federal), MCB (low research), AEM (industry research)
      + men from institutes/medical schools outperform women >7% for acceptance
      + >20% in favor of men at: EC (fed researh), mSphere & JVI (Low research), AEM & MCB (industry research), JCM/JB/JCM (R2 institution)
    + rev_score_analysis.R
      + reviewers are more likely to suggest rejections for women as compared to men. No difference in revise decisions
      + rejection suggestions are greater for women not from the US than those at US institutions
      + men at R1, low research, medical schools & fed research are favored by reviewers over women
    + reviewer_gender_analysis.R
      + both female & male reviewers are more likely to recommend rejection for women. 
      + Male reviewers are more likely to accept papers from men
      + women reviewers are more likely to recommend acceptance for women from low research & federal institutions 
      + women reviewers are more likely to recommend acceptance for men from R1 & industry instutitions
      + men are more likely to recommend acceptance for women from R2 institutions & favor men from R1 & medical institutes
    + editor_gender_analysis.R
      + both men & women editors are more likely to reject women
      + men editors are more likely to make revise only & accept decisions for men
      + women editors highly favor men from medical schools, slightly favor men from R2 & industry
      + men editors favor men from R1, R2 and medical schools, slightly favor women from low & industry research

  + Notes for plots:
      + highlight instances of >5% difference between genders
      + standardize gender terms to m/w
      + make faceting headers prettier

#Discussion

  + future directions
      + 

#Data and Methods

###Data

All manuscripts handled by ASM journals (mBio...) that recieved an editorial decision between January 1st, 2012 and August 31st, 2018 were supplied as XML files by ASM's publishing platform, eJP. Data were extracted from the XML documents provided using R statistical software (version 3.4.4) and the `XML` package. Data manipulation was handled using the `tidyverse`, `lubridate`, and `xml2` packages for R. Variables of interest included: the manuscript number assigned to each submission, manuscript type (e.g., full length research, erratum, editorial), category (e.g., microbial ecology), related (previously submitted) manuscripts, versions submitted, dates (e.g., submission, decision), author data (e.g., first, last, and corresponding authorships, total number of authors), reviewer data (e.g., reviewer score, recommendation, editor decision), and person data (names, institutions, country) of the editors, authors, and reviewers. 

It is common practice at ASM journals for manuscripts whose reviewers recommend extensive experimental revisions be given a decision of "reject with resubmission encouraged". If resubmitted, the authors are asked to note the previous (related) manuscript and the resubmission is assigned a new manuscript number. Multiple related manuscripts were tracked together by generating a unique grouped manuscript number based on the recorded related manuscript numbers. This grouped manuscript number served multiple purposes including: tracking a single manuscript through multiple rejections or transfers between ASM journals and to avoid duplicate counts of the same authors. 

###Gender assignment

The gender assignment API genderize.io was used to predict an indivdual's gender based on their given names, and country where possible. The genderize.io platform uses data gathered from social media to predict gender based on given names with the option to include an associated language or country to enhance the odds of succesful prediction. Since all manuscripts are submitted in English, we matched each individuals country against the list of X country names accepted by genderize.io. Using the `GenderGuesser` package for R, all unique given names associated with an accepted country were submitted to the genderize.io API and any names returned without a predictive assignment of either male or female were resubmitted without an associated country. All predictive assignments of either male or female are returned with a probability match of 0.50 or greater. The predicted genders of all given names (with and without an associated country) whose probabilities were greater or equal to our arbitrary success cut off of 0.65 were used to assign predicted gender to the individuals in our dataset. Predicted genders were assigned to individuals in the following order: first names and country, first names, middle names and country, middle names. Individuals who were not assigned a gender are given values of "N/A" and referred to in the analysis as those with an "unknown" gender. The predicted (male/female) gender of editors and senior editors in our dataset was hand validated using Google. 

###Data analysis and presentation


#References
