---
title: "**Who are ASM Journals? A Gender-based Analysis**"
csl: mbio.csl
fontsize: 11pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
geometry: margin=1.0in
bibliography: references.bib
---

```{r setup, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(knitr)
library(kableExtra)
library(markdown)
library(rmarkdown)
library(formattable)
library(cowplot)
library(tidyverse)
library(rlang)

load("../submission/gender_analysis.RData")

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")

```

\vspace{35mm}

Running title: A gender-based analysis of ASM journals

\vspace{35mm}


Ada K. $Hagan^{1,2}$, Begüm D. Topçuoğlu${^2}$, Hazel Barton${^3}$, Patrick D. Schloss^2$\dagger$^

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

1\. Alliance SciComm & Consulting, Linden, MI

2\. Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI

3\. Department of Biology, University of Akron, Akron, OH



\newpage
\linenumbers


##Abstract

Despite 50% of biology Ph.D. graduates being women, the number of women that advance in academia decrease at each level (e.g. to postdoctoral positions and tenure-track positions). Recently, scientific societies and publishers have begun examining internal submissions data to evaluate representation of, or bias against, women in their peer review processes; however, representation and attitudes differ by scientific field and no studies to-date seem to have investigated academic publishing in the field of microbiology. Using manuscripts submitted between January 2012 and August 2018 to 13 journals published by the American Society for Microbiology (ASM), we describe the representation of women at ASM journals and the outcomes of their manuscripts. We find that senior women authors at ASM journals are underrepresented compared to global and society estimates of microbiology researchers. Additionally, manuscripts submitted by corresponding authors that were women, received more negative outcomes (e.g., editorial rejections, reviewer recommendations, and decisions after review), than those submitted by men. These negative outcomes were somewhat mediated by whether or not the corresponding author was based in the US or not, and by the institution for US-based authors. Nonetheless, the pattern for women corresponding authors to receive more negative outcomes for their submitted manuscripts indicates a pattern of gender-influenced editorial decisions. We conclude with suggestions to improve the representation of, and decrease bias against, women at ASM journals.

##Importance

Women have fought an ongoing battle to become researchers and experts viewed equivalent to their colleagues who are men. We evaluated the participation and success of women researchers at ASM journals to better understand their success in the field of microbiology. We found that women are underrepresented as senior scientists at ASM journals. This is due to a combination of both low submissions from senior women authors and increased rejection rates for women compared to men.

##Introduction

Evidence has accumulated over the decades that academic research has a representation problem. While at least 50% of biology Ph.D. graduates are women, the number of women in postdoctoral positions and tenure-track positions are less than 40 and 30%, respectively [@sheltzer_elite_2014]. Studies examining other metrics such as race and ethnicity found that in 2011, less than 27% of all science and engineering doctorates were awarded to underrepresented minorities, and less than 25% of associate professors identified as non-white [@NCES_condition_2011; @NCES_full-time_2011]. Predictably, the disparities increase with academic rank [@potvin_diversity_2018]. There have been many proposed reasons for these disparities, particularly with regards to women, which include biases in training and hiring, the impact of children on career trajectories, a lack of support for primary caregivers, a lack of recognition, lower perceived competency, and less productivity as measured by research publications [@moss-racusin_science_2012; @sheltzer_elite_2014; children/caregivers; @aakhus_gender_2018; @broderick_gender_2019; @BlairLoy2017; @symonds_gender_2006]. These issues do not act independent of each other, instead they accumulate for both individuals and the community, much as advantages do [@diprete_cumulative_2006; @thebaud_segregation_2018]. Accordingly, addressing these issues necessitates multi-level approaches from all institutions and members of the scientific community.

Scientific societies play an integral role in the formation and maintenance of scientific communities. They host conferences that provide a forum for knowledge exchange, networking, and opportunities for increased visibility as a researcher. Scientific societies also frequently publish the most reputable journals in their field, facilitating the peer review process to vet new research submissions [@Schloss2017]. Recently, scientific societies and publishers have begun examining internal submissions data to evaluate representation of, and bias against, women in their peer review processes. The American Geological Union found that while the acceptance rate of women-authored publications was greater than publications authored by men, women submitted fewer manuscripts than men and were used as reviewers only 20% of the time [@lerback_journals_2017], a factor influenced by the gender of the editor [@fox_editor_2016]. Several other studies have concluded that there is no significant bias against papers authored by women [@ceci_understanding_2011; @fox_editor_2016; @handley_examination_2015; @edwards_gender_2018; @buckley_is_2014]. Several recent reports of outcomes at publishers for ecology and evolution, physics, and chemistry journals have found that women-authored papers are less likely to have positive peer reviews and outcomes [@Murray400515; @fox_gender_2019; @Physics_2018; @RoyalChem_2019]. 
  
The representation of women scientists and gender attitudes differ by scientific field and no studies to-date have investigated academic publishing in the field of microbiology. The American Society for Microbiology (ASM) is one of the largest life science societies, with an average membership of 41,000 since 1990. In its mission statement, the ASM notes that it is "an inclusive organization, engaging with and responding to the needs of its diverse constituencies" and pledges to "address all members' needs through development and assessment of programs and services." One of these services is the publication of microbiology research through a suite of research and review journals. Between January 2012 and August 2018, ASM published 25,818 original research papers across 15 different journals: _Antimicrobial Agents and Chemotherapy_ (AAC), _Applied and Environmental Microbiology_ (AEM), _Clinical and Vaccine Immunology_ (CVI), _Clinical Microbiology Reviews_ (CMR), _Eukaryotic Cell_ (EC), _Infection and Immunity_ (IAI), _Journal of Bacteriology_ (JB), _Journal of Clinical Microbiology_ (JCM), _Journal of Virology_ (JVI), _mBio_, _Microbiology and Molecular Biology Reviews_ (MMBR), _Genome Announcements_ (GA, now _Microbiology Resource Annoucements_), _Molecular and Cellular Biology_ (MCB), _mSphere_, and _mSystems_. The goal of this research study was to describe the population of ASM journals through the gender-based representation of authors, reviewers, and editors and the associated peer review outcomes. 
  
##Results

Over 100,000 manuscript records were obtained for the period between January 2012 and August 2018 (Fig. 1). Each of these were evaluated by reviewers and/or editors, leading to multiple possible outcomes. Manuscripts may be immediately rejected by editors instead of being sent to peer review, often due to issues of scope or quality. These were defined as editorial rejections and identified as manuscripts rejected without review. Alternately, editors send many manuscripts out for review by two or three experts in the field selected from a list of potential reviewers suggested by the authors and/or editors. Reviewers give feedback to the authors and editor, who decides whether the manuscript in question should be accepted, rejected, or sent back for revision. At ASM journals, manuscripts with suggested revisions that are expected to take more than 30 days are rejected, but generally encouraged to resubmit. If resubmitted, the authors are asked to note the previous (related) manuscript and the re-submission is assigned a new manuscript number. Multiple related manuscripts were tracked together by generating a unique grouped manuscript number based on the recorded related manuscript numbers. This grouped manuscript number served dual purposes of tracking a single manuscript through multiple rejections and avoiding duplicate counts of authors for a single manuscript. After eliminating non-primary research manuscripts and linking records for resubmitted manuscripts, there were 79,189 unique manuscripts processed (Fig. 1). 

We assigned genders to both peer review gatekeepers (e.g., editor-in-chief, editors, reviewers) and authors on the manuscripts evaluated during this time period using a classification algorithm (Supp Text). We recognize that biological sex (male/female) is not always equivalent to the gender that an individual presents as (man/woman), which is also distinct from the gender(s) that an individual may self-identify as. For the purposes of this manuscript, we choose to focus on the presenting gender (man/woman/unknown) based on their first names (and appearance for editors), as this is information is what reviewers and editors also have available. Author genders were assigned using a social media-informed predictive algorithm with stringent criteria and validation process (see methods). In addition to identifying journal participants as men or women, this method of gender assignment resulted in a category of individuals whose gender could not be reliably predicted (i.e., unknown). Among the individuals classified as either men or women, the sensitivity and specificity were 0.97 when validated against a curated set of authors and editors (Supp Text).

In the interest of transparency, we include those individuals whose names don't allow a high degree of confidence for gender assignment in the "unknown" category of our analysis, which is shown in many of the plots depicting representation of the population, but are not included in the comparison of manuscript outcomes.


**Men dominate as gatekeepers and senior authors.** We first evaluated the representation of men and women who were gatekeepers during the study period. Each journal is led by an editor-in-chief (EIC) who manages journal scope and quality standards through a board of editors with field expertise that, in turn, handle the peer review process. There were `r num_eic` EICs, `r prop_eic[[1,3]]`% of which were women. Two journals, EC and CVI were retired during the period under study. Four years before retirement, the leadership of CVI transferred from a man EIC, to a woman while JVI has had a woman as EIC since 2012. The total number of editors at all ASM journals combined over the duration of our study (senior editors and editors pooled) was `r num_ed`, `r prop_ed[[1,3]]`% of which were women.  

Over 40% of both men and women editors were from US-based R1 institutions, defined as doctoral universities with very high research activity, with non-US institutions, and U.S. medical schools or research institutions supplying the next largest proportions of editors (Fig. 2A). Since 2012, there was a slow trend toward equivalent gender representation among editors (Fig. 2B). The trends for each journal studied vary considerably, though most have slow trends toward parity (Fig. S1). CVI and _mSphere_ were the only ASM journals to have accomplished equivalent representation of both genders, with CVI having a greater proportion of women editors than men before it was retired. EC was the only journal with an increasing parity gap.

Altogether, `r num_rev` reviewers submitted reviews. Of those to whom we could assign a gender, `r prop_rev[[1,3]]`% were women. The greatest proportion of reviewers (over 50% of all groups) came from non-US institutions, while R1 institutions supplied the next largest cohort of reviewers (Fig. 2C). The proportions of each gender were steady among reviewers at ASM journals (Fig. 2D) and representative of both the suggested reviewers at all journals combined, and the actual reviewer proportions at most journals (Fig. S2). 

**Editorial workloads were not proportionate** Across all journals, men handled a slightly greater proportion of manuscripts than women, relative to their respective editorial representations (Fig. 3A). This trend continued across most journals with varying degrees of difference between workload and representation (Fig. S1). For instance, at _mSphere_, workload and proportions were identical; however, CVI, _mBio_, and JVI, each had periods at which the workload for women editors was much higher than their representation, with corresponding decreases in the workload of men. In the years preceding its retirement, the representation of women at CVI increased, decreasing the gap in editorial workload. However, representations and relative workloads for men and women editors at JVI have held steady over time, while the proportionate workload for women at _mBio_ has increased. 

  The median number of manuscripts reviewed by men, women, and unknown individuals was `r med_rev[[1,2]]`, for all groups. Half of those in the men, women, or unknown gender groups reviewed between one and `r med_rev[[2,3]]+1`, `r med_rev[[1,3]]+1`, or `r med_rev[[3,3]]+1` manuscripts each, respectively (Fig. 3B). Conversely, `r single_reviewer[[1,4]]`% of men, `r single_reviewer[[2,4]]`% of women, and `r single_reviewer[[3,4]]`% of unknown gendered reviewers reviewed only one manuscript. Reviewers of all genders accepted fewer requests to review from women editors (average of `r round(avg_f_ed_resp[[1,2]], digits = 2)`%) than from men (average of `r avg_m_ed_resp[[1,2]]`%) (Fig. 3C). Reviewers were less likely to respond to women editors than men (no response rate averages of `r avg_f_ed_resp[[2,2]]` and `r round(avg_m_ed_resp[[2,2]], digits = 2)`%, respectively). Editors of both genders contacted reviewers from all three gender groups at equivalent proportions, though women editors contact an average of `r med_contacted[[1,2]]`% of suggested reviewers, while men contact `r med_contacted[[2,2]]`% on average. 
     
```{r author_gender}
source("../code/representation/log_reg_prop_fem_auth.R") #prop.fem logistic regression weight & auc to determine corresponding author gender
```     
**Women were underrepresented as authors**  Globally, microbiology researchers are 60% men and 40% women [@allagnat_gender_2017]. In September 2018, `r get_percent(6214, 16193)`% of ASM members who reported their gender were women. We wanted to determine if these proportions were similar for authors at ASM journals and to understand the distribution of each gender among submitted manuscripts and published papers. We began by describing author institutions by gender. Over 60% of submitting senior authors (last or corresponding) were from non-US institutions, followed by about 20% from R1 institutions (Fig. 4A). The proportions of all men and women authors at ASM have decreased over time at equivalent rates, as the proportion of unknown gendered authors has increased, with a ratio of men to women authors of 4:3 (i.e., 57% men) (Fig. 4B).

In the field of microbiology, order of authorship on manuscripts signals the type and magnitude of contributions to the finished product with first and last authorship being the most prestigious. First authors are generally trainees (e.g., students or post-docs) or early career researchers responsible for the bulk of the project, while last authors are lead investigators, supplying conceptual guidance and resources to complete the project. Middle authors are generally responsible for technical analyses and methods. Any author can also be a corresponding author, which we identified as the individual responsible for communicating with publishing staff during peer review (as opposed to an author to whom readers direct questions).

The proportion of manuscripts submitted with men and women first authors have remained constant with averages of `r f_authors_avg_prop[[4,3]]` and `r f_authors_avg_prop[[3,3]]` percent, respectively (Fig. 4C, dashed). Their respective proportions of average published papers were nearly identical at `r f_authors_avg_prop[[2,3]]`% for men and `r f_authors_avg_prop[[1,3]]`% for women. The proportion of submitted manuscripts with men corresponding authors has remained steady at an average of `r c_authors_avg_prop[[4,3]]`% and the proportion with women corresponding authors was at `r c_authors_avg_prop[[3,3]]`% (Fig. 4D, dashed). Both men and women corresponding authors have a greater proportion of papers published than manuscripts submitted, just as manuscripts with corresponding authors of unknown gender were rejected at a higher rate. The difference between submitted manuscripts and published papers where men were corresponding authors is `r c_authors_avg_prop[[2,3]] - c_authors_avg_prop[[4,3]]`, but only `r c_authors_avg_prop[[1,3]] - c_authors_avg_prop[[3,3]]` where women were corresponding authors. This trend is similar for middle and last authors (Fig. S3). 

Of `r sum_m_corres` multi-author manuscripts submitted by men corresponding authors, `r get_percent(m_corres_prop_fem[[1,2]], sum_m_corres)`% had zero women authors. In contrast, `r most_w_auth_f_corres` (`r get_percent(most_w_auth_f_corres, sum_f_corres)`%) of manuscripts submitted by women corresponding authors had a majority of the authors as women, exceeding those submitted by men corresponding authors in both the number (`r most_w_auth_m_corres`) and percent (`r get_percent(most_w_auth_m_corres, sum_m_corres)`) of submissions. Additionally, the proportion of women authors decreased as the number of authors increases (Fig. S4). Men submitted `r sing_auth_num[[2,2]]` single-authored manuscripts while `r sing_auth_num[[1,2]]` single-authored manuscripts were submitted by women.

We hypothesized that we would be able to predict the gender of the corresponding author from the manuscript metadata. We trained a logistic regression model to predict the gender of the corresponding author using whether or not the corresponding author's institution was in the U.S., the total number of authors, the proportion of authors that were women, whether or not the paper was published, the gender of senior editors and editors, the number of revisions, and whether or not the manuscript was editorially rejected as variables to train our model. We measure the model's performance using the <u>a</u>rea <u>u</u>nder the <u>r</u>eceiver <u>o</u>perating <u>c</u>haracteristic curve (AUROC). The AUROC value is a predictive performance metric that ranges from 0.0, where the model's predictions are  completely wrong, to 1.0, where the model perfectly distinguishes between outcomes. A value of 0.5 indicates the model did not perform better than random. The median AUROC value of our model to predict corresponding author gender was 0.7. The variable with the largest weight (i.e., the most predictive value), in our model was the proportion of women authors. These results indicate that manuscript submission data was capable of predicting the gender of the corresponding author, but that prediction is primarily driven by what percentage of authors are women.

As described above, first authors were slightly more likely to be women (`r f_authors_avg_prop[[3,3]]` vs `r f_authors_avg_prop[[4,3]]`), but corresponding authors were more likely to be men (`r c_authors_avg_prop[[3,3]]` vs `r c_authors_avg_prop[[4,3]]`). A concern is that if authors are not retained so they transition from junior to senior status, they are also left out of the gatekeeping roles. Since authorship conventions indicate that last and corresponding authors are senior, we combined both first and middle authors into the "junior" author role. There were `r percent_retent[[1,4]]` women who participated as junior authors (first/middle) at ASM journals. Of those junior authors who were women, `r percent_retent[[2,5]]`% also participated as senior authors (last/corresponding), `r percent_retent[[3,5]]`% were suggested reviewers and `r percent_retent[[4,5]]`% participated as reviewers. `r percent_retent[[5,5]]`% of women junior authors were also editors at ASM journals. For men, there were a total of `r percent_retent[[6,4]]` junior authors, where `r percent_retent[[7,5]]`% also participated as senior authors, `r percent_retent[[8,5]]`% considered as reviewers, and `r percent_retent[[9,5]]`% actually reviewed. `r percent_retent[[10,5]]` of men junior authors were also editors at ASM journals. Overall, women were half as likely to move to more prestigious (e.g., senior author, gatekeeper) roles in peer review than men.

```{r difference_rejections} 
source("../code/bias/log_reg_gend_rej.R") #logistic regression auc 
#ASM-rej-rate comes from Figure_5.R
```
**Manuscripts submitted by women have more negative outcomes than those submitted by men.** To better understand the differences between published and submitted proportions for men and women authors (Fig. 4CD, Fig. S3), we compared the rejection rates of men and women at each author stage (first, middle, corresponding, and last). Middle authors were rejected at similar rates for men and women, a `r ASM_rej_rate[4,5]` percentage point difference across all journals; however, manuscripts with senior women authors were rejected more frequently than those authored by men with `r ASM_rej_rate[1,5]` and `r ASM_rej_rate[3,5]` percentage point differences for corresponding and last authors, respectively (Fig. 5A, vertical line).  There were several instances where the overall trend was repeated at the journal level (e.g., AAC, IAI, JB, _mBio_, MCB). The greatest differences were observed when comparing the outcome of corresponding authors by gender, so we used this sub-population to further examine the difference in manuscript acceptance and rejection rates between men and women.

We next compared the rejection rates for men and women corresponding authors after two bottlenecks, initial review by the editor and the first round of peer review. Manuscripts authored by women were editorially rejected as much as 12 percentage points more often than those authored by men (Fig. 5B). The percentage point difference at all ASM journals combined was `r fig4_ASM_ed_rej[[1,4]]` (vertical line). MCB and _mBio_, had the most  extreme percentage point differences. Manuscripts authored by men and women were equally likely to be accepted after the first round of review (Fig. 5C, right panel). However, women-authored papers were rejected more often (left panel) while men-authored papers were more often given revision decisions (center panel). JB, AAC, and MCB had the most extreme differences rejection and revision decisions, (`r fig4_ASM_dec[[1,4]]` and `r fig4_ASM_dec[2,4]` respectively; Fig. 5C vertical lines). Percentage point differences were not correlated with journal prestige as measured by impact factors (R${^2}$ = -0.022, P = 0.787). 

In addition to manuscript decisions, other disparate outcomes may occur during the peer review process [@erin_hengel_publishing_2017]. To determine whether accepted women-authored manuscripts spent more time between being submitted and being ready for publication, we compared the number of revisions, days spent in the ASM peer review system, and the number of days from submission to being ready for publication to those authored by men. Manuscripts authored by women took slightly longer (from submission to ready for publication) than men at some journals (_mSphere_, _mBio_, _mSystems_, CVI, JB, JCM, AEM) despite spending similar amounts of time in the ASM journal peer review system (Fig. S5), and having the same median number of revisions prior to acceptance (Median = 2, IQR = 0). 

To understand how gatekeeper (editor/reviewer) genders influence decisions (e.g., Fig. 5C), we grouped editor decisions and reviewer suggestions according to the gatekeeper gender. Both men and women editors rejected proportionally more women-authored papers, however the difference in decisions were more extreme for men-edited manuscripts (Fig. 6A). Reviewers were more likely to suggest rejection for women-authored manuscripts as compared to men, although a minimal difference in revise recommendations was observed (Fig. 6B). Both men and women reviewers recommended rejection more often for women-authored manuscripts although men recommended acceptance more often for men-authored manuscripts (Fig. 6C).

To evaluate if gender played a role in manuscript editorial decisions, we trained a L2-regularized logistic regression model to predict whether or not a manuscript was reviewed (i.e. editorially rejected or not). We used the genders of senior editor, editor, and corresponding author, as well as the proportion of authors that were women as variables to train the model. The median AUROC value was `r med_auc_gend`, which indicated that editorial decisions were not random, however, the AUROC value is relatively low indicating that there are factors other than the variables we included in our model that influence editorial decisions. 

```{r other_factors} 
#source("../code/bias/log_reg_US_rej.R") #logistic regression auc 
#ASM-rej-rate comes from Figure_5.R
```
**Multiple factors contribute to the overperformance of men.** The association between gender and manuscript decision could be attributed to gender bias by journal gatekeepers, however, there are other types of bias that may contribute to, or obscure, overt gender bias; for instance, a recent evaluation of peer-review outcomes at _eLife_ found evidence of preference for research submitted by authors from their own country or region [@Murray400515]. Other studies have documented prestige bias, where men are over-represented in more prestigious (i.e., more respected and competent) programs [@weeden_degrees_2017]. It is therefore possible, that what seems to be gender bias could be geographic or prestige bias interacting with the increased proportion of women submitting from outside the US or at lower prestige institutions (e.g., low research institutions) (Fig. 4A).

To quantify how these factors affect manuscript decisions, we next looked at the outcome of manuscripts submitted only by corresponding authors at US institutions (31% submitted by women in the US vs 36% non-US), because these institutions represented the majority of manuscripts and are classified by the Carnegie Foundation. When only considering US-based authors, the difference in percentage points for editorial rejections increased from  `r fig4_ASM_ed_rej[[1,4]]` to `r fig6_ASM_ed_rej[[1,3]]` (Fig. 7A). Similarly, the difference in percentage point trends of decisions after review for US-based authors mirror those seen for all corresponding authors at the journal level (Fig. 7B). The over-performance of women in rejection decisions increased from `r fig4_ASM_dec[[1,4]]` to `r US_j_dec[[1,4]]`, and the over-performance of men in revise only decisions decreased from `r fig4_ASM_dec[[2,4]]` to `r US_j_dec[[2,4]]` (Fig. 7B). The rate of accept decisions changed from `r fig6_ASM_ed_rej[[1,3]]` to `r US_j_dec[[3,4]]` after restricting the analysis to US-based authors. These results suggest that the country of origin (e.g, US versus not) accounted for some gender bias, particularly for editorial rejections.

To address institution-based prestige bias, we split the US-based corresponding authors according to the type of institution they were affiliated with and re-evaluated the differences for men and women. Editorial rejections occurred  most often for women from medical schools or institutes, followed by those from R2 institutions, 28% and 34% of manuscripts, respectively (Fig. 7C). Manuscripts submitted by men from medical schools or institutes were accepted proportionally more often than those submitted by women. The difference for corresponding authors from medical schools or institutes seems to be spread across most ASM journals, while the editorial rejection of papers from R2 institutions seems to be driven primarily by JCM (Fig. S6AB). 

To evaluate if these factors affect manuscript decisions, we trained a L2-regularized regression model to predict if a manuscript was editorially rejected or not, using the variables: origin (US vs non), institution (US institution type), number of authors, proportion of authors that were women, and the genders of both gatekeepers and authors. The model had a median AUROC value of `r med_auc_US` which indicated that there was a non-random interaction between these factors and editorial decisions. Manuscripts from authors at U.S. "other" institutions, men EICs, men that are corresponding authors from "other" U.S. institutions, and women from medical schools and institutes were more associated with editorial rejections (Fig. S6C). Conversely, manuscripts from R1 institutions, authors from the U.S., EICs that were women, and the number of authors were more likely to be associated with review (Fig. S6C). These results confirm that the country of origin and class of institutions impact decisions in a non-random manner, though not as much as gender.

A final possibility is that the type of research pursued by men as opposed to women may impact manuscript outcomes. Black women philosophers and physicists have described systems of "epistemic oppression", or devaluation of non-traditional knowledge and theories, in their fields [@Dotson2012; @Dotson2014; @PrescodWeinstein2020]. While orginally focused on bias against Black women---the intersection of two historically marginalized identities---the concept that those in the historical center of a field might be skeptical of non-traditional research likely applies elsewhere. This phenomenon has recently been observed in the biomedical sciences where NIH proposals focusing on women's reproductive health were the least likely to be funded [@Hoppe2019]. To briefly explore this possibility at ASM journals, we looked at the editorial rejection rates of each research category at five, longstanding ASM journals: AAC, AEM, IAI, JVI, and JCM (Fig. S7). Each journal focuses on a different facet of microbiology or immunology, making the results difficult to compare. However, there were a couple observations we would like to note. One category, Biologic Response Modifier (e.g., immunotherapy) at AAC, had extreme differences for both editorial rejections and rejections after review, about -30 and -40, respectively. While that category had a comparatively low number of submissions (44), 43% were from women (Fig. S7A). Despite the difference in research fields between journals, there was one category, Mycology, which was the same at two journals, AEM and JCM. At both journals, men overperformed, relative to women in this category. At AEM, there were 73 Mycology submissions, 44% from women authors that had a difference of almost -20 for editorial rejection outcomes and -10 for rejections after review (Fig. S7B). Conversely, JCM had 587 Mycology submissions with a similar submission rate of 39% from women authors (Fig. S7D). Differences between outcomes were almost -10 for editorial rejections and -12 for rejections after review. 

We next asked if the amount of women in a particular category was related to manuscript outcomes. There was no correlation between difference in editorial rejection performance by category and the percent of women that were authors (R${^2}$ = -0.003, P = 0.363) or editors (R${^2}$ = -0.018, P = 0.765). Neither was there a correlation between the percent of women authors and editors in journal categories (R${^2}$ = -0.007, P = 0.682). These data suggest the possibility of persistant (e.g., by editors and reviewers) bias against women in particular fields (e.g., Mycology), though it does not seem to relate to the number of women publishing in those fields.  

##Discussion

We described the representation of men and women at ASM journals between January 2012 and August 2018 and compared editorial outcomes according to the authors' gender. Women were consistently under-represented (30% or less in all levels of the peer review process) excluding first authors, where women represented about 50% of authors where we could assign a gender (Fig. 2/4). Women and men editors had proportionate workloads across all ASM journals combined, but those workloads were disproportionate at the journal level and the overburdened gender varied according to the journal (Fig. 3/S1). Additionally, manuscripts submitted by corresponding authors that were women, received more negative outcomes (e.g., editorial rejections), than those submitted by men (Fig. 5/6). These negative outcomes were somewhat mediated by whether the corresponding author was based in the US, the type of institution for US-based authors, and the research category (Fig. 7/S7). However, the trend for women corresponding authors to receive more negative outcomes held, indicating a pattern of gender-influenced editorial decisions regardless of journal prestige (as determined by impact factor).

The proportion of women as first authors is higher than data obtained globally and from self-reported ASM membership data, which in turn was higher than the proportion of senior women authors at ASM journals. Only half as many women who were junior authors at ASM journals were also senior authors when compared to men, and the representation of women decreased as the prestige (e.g., reviewer, editor) increased. These trends are consistent with representation of senior women in academic biological sciences and the observation that women are more likely to leave academia during the transition from postdoc to investigator [@Martinez2007]. These data indicate that microbiology (as represented by ASM journals) is not exempt from the issues that limit the retention of women through academic ranks. 

How to define representation and determine what the leadership should look like are recurring questions in STEM. Ideally, the representation for men and women corresponding authors, reviewers, and editors would reflect the number of Ph.D.s awarded (about 50% each, when considered on a binary spectrum). We argue that the goal should depend on the workload and visibility of the position(s). Since high visibility positions (e.g., editor, EIC) are filled by a smaller number of individuals that are responsible for recruiting more individuals into leadership, filling these positions should be done aspirationally (i.e., 50% should be women if the goal were an aspirational leadership). This allows greater visibility for women as experts, expansion of the potential reviewer network, and recruitment into those positions [@debarre_gender_2018; @sardelis_not_2016; @casadevall_presence_2014]. Conversely, lower visibility positions (e.g., reviewers) require a greater number of individuals and should thus be representational of the field to avoid overburdening the minority population (i.e., since 23.5% of corresponding authors to ASM journals are women, then 20-25% of reviewers should be women). Balancing the workload is particularly important given the literature indicating that women faculty have higher institutional service loads than their counterparts who are men [@guarino_faculty_2017].

In contrast to institutional service, the editing workload at ASM journals seems to be predominantly borne by men. A possible explanation for the difference in gatekeeper representation and editor workloads is that women are more likely to study non-traditional research fields [@Dotson2012; @Dotson2014; @PrescodWeinstein2020]. Their separation from the traditional center of a field decreases their perceived competency, which could result in research typecasting and lower manuscript handling responsibilities. Any of these situations can be compounded by the increased proportion of potential reviewers that either do not accept, or do not respond to, requests to review from women editors. This increases the proportion of reviewers that women editors must contact, adding additional time and work to their editorial burdens. Three journals, _mBio_, CVI, and JVI are exceptions with regards to editorial workloads. At these journals, the editorial workloads of women exceeds their representation. A possible explanation for CVI and JVI is that both of these journals have been led by women EICs. Alternately, the tendency for reviewers to reject requests to review from editors that are women, extends to editors that are men who may reject requests to handle manuscripts from EICs that are women. Our data differ from those of Fox, Burns, and Meyer who found that the gender of the editor influenced the gender of the contacted reviewers, but supports findings that women editors contact more reviewers than men [@fox_editor_2016; @gilbert_is_1994]. 

Previous research suggests that women who collaborate with other women receive less credit for these publications than when they collaborate with men [@wiedman_rewarding_2019], and that women are more likely to yield corresponding authorship to colleagues that are men [@fox_gender_2019]. Our data support findings of previous studies that women are more likely to collaborate with other women [@holman_researchers_2019; @RoyalChem_2019]. In our linear regression models, the number of authors on a manuscript was the largest contributor to avoiding editorial rejections, suggesting that highly collaborative research is preferred by editors, an observation supported by the increase of citations with author count [@fox_citations_2016]. It is, therefore, disturbing that when the number of authors exceeded 30 on a manuscript (N=59), the proportion of individuals predicted to be women was always below 51%, despite equivalent numbers of trainees in the biological sciences (Fig. S4). Additionally, while women corresponding authors submit fewer manuscripts, more of them (both numerically and proportionally), have a majority of women co-authors, compared to those submitted by men corresponding authors. Thus the gender-based segregation of collaborations at ASM journals likely has had consequences in pay and promotion for women participating at ASM journals and could be a factor in the decreased retention of senior women. Additionally, the under representation of women as corresponding authors in publication at ASM journals may have negative consequences for their careers and microbiology, since their role as senior author can impact their status in the field. Buckley et al, suggested that being selected as a reviewer increases visibility of a researcher, which has a direct and significant impact on salary [@buckley_is_2014]. Therefore, the under representation of women as senior authors and reviewers likely hampers their career progression and even their desire to progress since status in the peer review process also signals adoption of the researcher into the scientific community [@buckley_is_2014]. Retention of women in science is important to the progress of microbiology since less diversity in science limits the diversity of perspectives and approaches, thus stunting the search for knowledge. In addition to boosting productivity and knowledge, more diverse and equitable organizations are more inclusive and supportive for all members [@potvin_diversity_2018]. 

Whether academic research journals support women has been the topic of many papers, which note the lack of women authors publishing relative to men in high impact journals [@berg_examining_2019; @conley_call_2012-1; @bendels_gender_2018; @Shen275362]. However, submissions data is required to determine if the lack of representation is due to low submissions or bias during peer review. We have shown that there is a disparity in submissions from senior women in microbiology compared to men, but this does not fully account for the difference in publications by men and women corresponding authors at ASM journals. This is due to a consistent trend favoring positive outcomes for manuscripts submitted by corresponding authors that are men. Manuscripts submitted by corresponding authors who are women are editorially rejected at greater rates, and gatekeepers of both genders favor revisions for manuscripts authored by men but rejection for those authored by women. Neither geographic (e.g., US or not) nor institution type can fully account for the gender-based bias observed. Instead, the presence of bias favoring men over women from U.S. R1 institutions and medical schools and institutes suggests that the bias persists even in environments with generally excellent resources and infrastructure for research. The scientific endeavor and peer review system select for decisions that are often based on the assumption that scientists are objective, impartial experts. As a result, scientists who believe themselves immune to biases are making decisions that rely on biases to speed the process. The types of biases at play and their potential roles in peer review are well documented [@Kaatz2014; @Carnes2005]. For instance, previous studies document that a greater burden of proof is required for women to achieve similar competency as men and that women are less likely to self-promote (and are penalized if they do) [@BlairLoy2017; @babcock_women_2003; @MILLER1992]. These might lead women to be more conservative in manuscript submissions, making the observed bias even more concerning. 

Even if a gatekeeper does not know the corresponding author or their gender, there remain ample avenues for implicit bias during peer review. The stricter standard of competency has led women to adopt different writing styles from men, resulting in manuscripts with increased explanations, detail, and readability than those authored by men [@Kolev2019; @erin_hengel_publishing_2017]. These differences in writing can act as subtle cues to the author's gender. Additionally, significant time, funds, and staff are required to be competitive in highly active fields (e.g., _Clostridium difficile_, HIV), but women are often at a disadvantage for these resources due to the cumulative affects of bias [@diprete_cumulative_2006; @thebaud_segregation_2018]. As a result, corresponding authors that are women may be spending their resources in research fields where competition impacts are mitigated and/or on topics that are historically understudied. This has the disadvantage of further decreasing perceived competency of these women scientists to those studying historical field(s) with precedence [@Dotson2012; @Dotson2014; @PrescodWeinstein2020]. Alternatively, non-traditional research may be seen as less impactful, leading to poorer outcomes [@Hoppe2019]. These possibilities are reflected in our data, since while the number of revisions before publication is identical for both men and women, manuscripts authored by women have increased rejection rates and time spent on revision. This suggests that manuscripts submitted by women receive more involved critiques (i.e., work) from reviewers and/or their competency to complete revisions within the prescribed 30 days is doubted, when compared to men. Women may also feel they need to do more to meet expectations leading to longer periods between a decision and resubmission. Finally, our data show a penalty for women researching mycology (Fig. S7). Despite being among the most deadly communicable diseases in 2016 (along with tuberculosis and diarrheal diseases), mycology is an underserved, and underfunded, field in microbiology that has historically been considered unimportant [@Head2014; @WHO_2018, @ASM_2019].

Few papers have found disparities between the rejection rates of men and women [@RoyalChem_2019; @Physics_2018; @fox_gender_2019]. To our knowledge, this is the first paper to examine this issue in the field of microbiology. A limitation to our methodology is the use of an algorithm to assign gender by first names. This method left us with a category of unknown gendered individuals and the gender of an individual maybe interpreted differently according to the reader (e.g., Kim is predominately a woman's name in the U.S., but likely a man's name in other cultures). The increase in unknown gendered authors corresponds to an increase in submissions to ASM journals from Asian countries, particularly China. Anecdotally, most editorial rejections are poor quality papers from Asia and our method has low performance on non-gendered languages from this region (see Supplemental Text). As a result, many manuscripts from Asia were excluded from the analysis on decision outcomes, increasing our confidence that the trends observed are gender-based. Another concern might be the small effect size observed in many analyses. Nonetheless, the consistency of decisions to benefit men corresponding authors over women, across all journals included in this study, in addition to accumulated literature to-date, confirms that this descriptive study is highly relevant for the ASM as a society. Our findings offer opportunities to address gendered representation in microbiology and systemic barriers to peer review at our journals.

All parties have an opportunity and obligation to advance underrepresented groups in science [@potvin_diversity_2018]. We suggest the development of a visible mission, vision, or other commitment to equity and inclusion that includes a non-discrimination clause regarding decisions made by editors and editors-in-chief. This non-discrimination clause would be backed by a specific protocol for the reporting of, and responding to, instances of discrimination and harassment. Second, society journals should begin collecting additional data about authors and gatekeepers (e.g., race, ethnicity, gender identity, and disabilities). Such author data should not be readily available to journal gatekeepers, but instead kept in a dis-aggregated manner that allows for public presentation to track success of inclusive measures and maintain accountability in a more accurate manner than this report. Third, society journals can implement mechanisms to explicitly provide support for women and other minority groups, reward inclusive behavior by gatekeepers, nominate more women to leadership positions and recruit manuscripts from sub-fields that are more likely to attract women and other minorities [@Hoppe2019]. Gatekeepers, and authors, can help advance women (and other minority groups) within the peer review system by changing how they select experts in their field. For instance, authors can suggest more women as reviewers using "Diversify" resources, while reviewers can agree to review for women editors more often [@Hagan_2019]. Editors can rely more on manuscript reference lists and data base searches than personal knowledge, and journals can improve the interactivity and functionality of the peer review selection software [@fox_gender_2016]. Growing evidence suggests that representation problems in STEM are due to retention rather than recruitment. We need to align journal practices to foster the retention of women and other minority groups.

Addressing bias (gender, geographic, prestige, or otherwise) during peer review process is a more difficult challenge, since it is partially the result of accumulated disadvantages and the actions resulting from implicit biases. Most approaches to overcoming these issues focus on choices made by individuals, such as double-blinded reviews and implicit bias training, but these cannot fully remedy the effects of bias and may even worsen outcomes [@cox_case_2018; @applebaum_remediating_2019]. Broadly, peer review is a nebulous process with expectations and outcomes that vary considerably, even within a single journal. Academic writing courses suffer similar issues and have sought to remedy them through the use of rubrics. When implemented correctly, rubrics can reduce bias during evaluation and enhance the evaluation process for both the evaluator and the evaluatee [@Holmes2011;@Malouff2016;@Reddy2010;@Rezaei2010]. We argue that rubrics could be implemented in the peer review process to focus reviewer comments, clarify editorial decisions, and improve the author experience. Such rubrics should increase the emphasis on **solid** research, as opposed to novel or "impactful" research, the latter of which is a highly subjective measure [@Casadevall2014; @Casadevall2015]. This might also serve to change the overall attitude toward replicatitive research and negative results, thus bolstering the field through reproduciblity. We also argue that reconsidering journal scope and expanding editorial boards might help address structural barriers bias against women (and other minorities) in peer review. Expanding journal scope and adding more handling editors would improve the breadth of research published, thus providing a home for more non-traditional and underserved research fields (the case at _mSphere_ with an increased pool of editors). Implementing these steps to decrease bias---review rubrics, increased focus on solid research, expansion of journal scopes, and editorial boards---will also standardize competency principles for researchers at ASM journals and improve microbiology as a whole. 

We hope that this report will help to counteract the bias against women through the transparency of accountability that provides our gatekeepers with the evidence of bias. However, the adaptation of women (and other minority groups) to bias (e.g., area of research and communication styles), make it impossible to level the playing field at the individual level. Instead, we must commit to changing the fundamental structure and goals of peer review to minimize bias.

##Data and Methods

**Data** All manuscripts handled by ASM journals (e.g., _mBio_, _Journal of Virology_) that received an editorial decision between January 1, 2012 and August 31, 2018 were supplied as XML files by ASM's publishing platform, eJP. Data were extracted from the XML documents provided, manipulated, and visualized using R statistical software (version 3.4.4) and relevant packages [@R_software_2017; @wickham_tidyverse_2017; @duncan_xml_2018; @wickham_xml2_2018; @grolemund_dates_2011; @wickham_scales_2018; @neuwirth_rcolorbrewer_2014; @cowplot; @rlang; @formattable; @knitr_2018; @knitr_2014; @rmd_rstudio; @rmd_book; @markdown]. Variables of interest included: the manuscript number assigned to each submission, manuscript type (e.g., full length research, erratum, editorial), category (e.g., microbial ecology), related (i.e., previously submitted) manuscripts, number of versions submitted, dates (e.g., submission, decision), author data (e.g., first, last, and corresponding authorship, total number of authors), reviewer data (e.g., recommendation, editor decision), and personal data (names, institutions, country) of the editors, authors, and reviewers. For this analysis, only original, research-based manuscripts were included, e.g., long- and short-form research articles, New-Data Letters, Observations, Opinion/Hypothesis articles, and Fast-Track Communications. To help protect the confidentiality of peer review, names have been removed from all records, and identifying data (e.g., manuscript numbers, days of date), have been replaced with randomized values.

**Institution classification** To identify the communities represented, we used Carnegie classifications to group US-based academic institutions into R1 research (very high research activity), R2 research (high research activity), four-year medical schools, or low research (i.e., not R1, R2, or medical school) [@Carnegie2018]. Research institutes (e.g., Mayo Clinic, Cold Springs Harbor), industry (e.g., pharmaceutical), and federal (e.g., NIH, CDC) research groups were identified using the internet. Four-year medical schools and research institutions were grouped together since these typically share research prestige and have considerable resources to support research. Industry and federal research were grouped separately. The "Other" category represents uncategorized US institutions. Non-US institutions were a category on their own.

**Gender prediction and assignment** The gender assignment API genderize.io was used to predict an individual's gender based on their given names and country where possible. The genderize.io platform uses data gathered from social media to predict gender based on given names with the option to include an associated language or country to enhance the odds of successful prediction. Since all manuscripts are submitted in English, precluding language association for names with special characters, names were standardized to ASCII coding (e.g., "José" to "Jose"). We next matched each individuals' country against the list of 242 country names accepted by genderize.io. Using the GenderGuesser package for R [@caddigan_genderguesser], all unique given names associated with an accepted country were submitted to the genderize.io API and any names returned without a predictive assignment of either male or female were resubmitted without an associated country. The data returned include the name, predicted gender (male, female, no prediction), the probability of correct gender assignment (ranging from 0.5 to 1.0), and the number of instances the name and gender were associated together (1 or greater). The predicted genders of all given names (with and without an associated country) whose probabilities were greater or equal to a modified probability (pmod) of 0.85 were used to assign predicted genders (man/woman) to the individuals in our data set (see Supplemental Text). Predicted genders were assigned to individuals in the following order: first names and country, first names, middle names and country, middle names. The presenting gender (man/woman) of editors and senior editors in our data set was hand verified using Google where possible and the algorithm validated using published data [@broderick_gender_2019].

**Manuscript outcome analysis** To better visualize and understand the differences in outcomes according to author gender, we calculated the difference in percentage points between the proportion of that outcome for men and women. To correct for the disparity in the participation of women relative to men at ASM journals, all percentage point comparisons are made relative to the gender and population in question. For instance, the percentage point difference in acceptance rates is the acceptance rate for men minus the acceptance rate for women. A positive value indicates that men received the outcome more often than women, whereas a negative value indicates that women outperformed men in the given metric.  

**Logistic regression models** For the L2-regularized logistic regression models, we established modeling pipelines for a binary prediction task. First, we randomly split the data into training and test sets so that the training set consisted of 80% of the full data set while the test set was composed of the remaining 20% of the data. To maintain the distribution of the two model outcomes found with the full data set, we performed stratified splits. The training data was used to build the models and the test set was used for evaluating predictive performance. To build the models, we performed an internal five-fold cross-validation where we tuned the cost hyper-parameter, which determines the regularization strength where smaller values specify stronger regularization. This internal cross-validation was repeated 100 times. Then, we trained the full training data set with the selected hyper-parameter values and applied the model to the held-out data to evaluate the testing predictive performance of each model. The data-split, hyper-parameter selection, training and testing steps were repeated 25 times to get a reliable and robust reading of model performance. Models were trained using the machine learning wrapper caret package (v.6.0.81) in R (v.3.5.0) [@khun_caret_2018].

**Code and data availability** Anonymized data and code for all analysis steps, logistic regression pipeline, and an Rmarkdown version of this manuscript, is available at https://github.com/SchlossLab/Hagan_Gender_mBio_2019/

**Acknowledgements**
We would like to thank Nicole Broderick and Arturo Casadevall for providing their data set for genderize validation and acknowledge Arianna Miles-Jay and Joshua MA Stough for their comments.

A.K.H. was responsible for data aggregation, analysis, interpretation, and drafting the manuscript. B.T. completed the logistic regression models. A.K.H., H.B., and P.D.S. were involved with conceptual development. Funding and resources were provided by P.D.S. All authors contributed to the final manuscript. P.D.S. is Chair of ASM Journals and A.K.H. was ASM staff prior to publication of the analysis. B.T. and H.B. report no conflict of interest.

Funding and access to the data for this work were provided by the American Society for Microbiology. Early drafts were read by the ASM Journals Committee with minimal influence on content or interpretation.

\newpage
\includegraphics{population_diagram.png}
**Figure 1. Overview of manuscript outcomes.** Over 100,000 manuscript records were obtained for the period between January 2012 and August 2018. After eliminating non-primary research manuscripts and linking records for resubmitted manuscripts, we processed 79,189 unique manuscripts. The median number of versions was 1 (iqr=0-2) with a median of 6 (iqr=1-11) authors per manuscript. As of August 2018, 34,196 of these were published. Revisions were requested for 24,016 manuscripts and 53,436 manuscripts were rejected at their first submission. The number of individuals (e.g., author, editor, reviewer) involved in each category of manuscript decision are indicated in the colored boxes: women (orange), men (blue), and unknown (gray). *A small number were given revise (242) or acceptance (1094) decisions without review. 

\newpage
\includegraphics{Figure_1.png}
**Figure 2. Gendered representation among gatekeepers.** Proportion of editors from (A) institution types and (B) over time. Editors and senior editors are pooled together. Proportion of reviewers from (C) institution types and (D) over time. (A,C) Each gender equals 100% when all institutions are summed.(B,D) Each individual was counted once per calendar year, proportions of each gender add to 100% per year. 

\newpage
\includegraphics{Figure_2.png}
**Figure 3. Gatekeeper workload and response to requests to review.** (A) Proportion of manuscript workloads by men and women editors, editorial rejections excluded. (B) Box plot comparison of all manuscripts, by reviewer gender. (C) The percent of reviewers by gender that accepted the opportunity to review, split according to the editor's gender. 

\newpage
\includegraphics{Figure_3.png}
**Figure 4. Author representation by gender.** The proportion of (A) men and women senior authors from each institution type, (B) men, women, and unknown authors from 2012 - 2018. Each individual was counted once per calendar year. The proportion of (C) first authors and (D) corresponding authors from 2012 - 2018 on submitted manuscripts (dashed) and published papers (solid).

\newpage
\includegraphics{Figure_4.png}
**Figure 5. Difference in rejection rates by corresponding author gender.** (A) The percent of manuscripts rejected by author gender and type (e.g., corresponding, first, last, middle) at any stage across all journals where 0 corresponds to equal rates of rejection. (B) The difference in percent editorial rejection rates for corresponding authors at each journal. (C) The difference in percentage points between each decision type for corresponding authors following the first peer review. Vertical lines indicate the difference value for all journals combined. Absence of a bar indicates parity.

\newpage
\includegraphics{Figure_5.png}
**Figure 6. Difference in decisions or recomendations according to the gatekeeper gender.** (A) Effect of editor gender on the difference in decisions following review. (B) Difference in percentage points for review recommendations and (C) how that is affected by reviewer gender. (A-C) All journals combined.

\newpage
\includegraphics{Figure_6.png}
**Figure 7. Impact of origin and U.S. institution type on manuscript decisions by gender.** Difference in percentage points for (A) editorial rejections and (B) following first review of manuscripts submitted by US-based corresponding authors. Vertical line indicates value for all ASM journals combined. (C) Difference in percentage points for acceptance and editorial rejections according to institution types and (D) acceptance decisions by to editor gender and institution types.  

\newpage
\includegraphics{Figure_S1.png}
Figure S1. The proportion of editors (solid line) and their workloads (dashed line) at each ASM journal from 2012 to 2018.

\newpage
\includegraphics{Figure_S2.png}
Figure S2. The proportion of (A) potential reviewers at all ASM journals combined, (B) reviewers at each ASM journal.

\newpage
\includegraphics{Figure_S3.png}
Figure S3. The proportion of all submitting (dashed line) and publishing (solid line) (A) middle and (B) last authors by gender at each ASM journal. 

\newpage
\includegraphics{Figure_S4.png}
Figure S4. The proportion of women authors on submitted manuscripts according to the number of authors and the gender of the corresponding author.

\newpage
\includegraphics{Figure_S5.png}
Figure S5. Comparison of time to final decision and impact by gender. The number days (A) between when a manuscript is initially submitted then finally published and (B) that a manuscript spends in the ASM peer review system. (C) The median number of versions between submission and publication.

\newpage
\includegraphics{Figure_S6.png}
Figure S6. Difference in A) editorial rejection and B) acceptance rates by journal and institution type. C) Difference in review recommendations by reviewer gender and author institution type. D) Median importance (black dot) of features affecting editorial rejections, and their range. Color of smaller dots (n=25) indicate the direction of the impact. 

\newpage
\includegraphics{Figure_S7.png}
Figure S7. Difference in editorial rejections by corresponding author gender and manuscript category at (A) AEM, (B) JCM, and (C) JVI.

\newpage
\includegraphics{Figure_S8.png}
Figure S8. (A) Equation for calculating negative bias by genderize. C indicates a country. (B) The negative impact of each country on the overall gender prediction of the full data-set. Number is the total number of names associated with each country.

\newpage
## References
