---
title: "Gender bias during peer review at ASM journals is mediated by author institution and gatekeeper gender"
author: "Ada Hagan, Begum Topcholou, Hazel Barton, Patrick Schloss"
date: "July 17, 2018"
output:
  html_document:
    df_print: paged
csl: asm.csl
bibliography: references.bib
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
library(markdown)
library(rmarkdown)
library(formattable)
library(cowplot)

source("../code/load_data.R")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

#Abstract

#Introduction

  + Gender disparity in stem
  
  + Studies evaluating gender in publications
      + Nature
      + JAMA
      + Glam journals
      + Science?
      + Ecology (Cho, 2014; Pettorelli, 2013; Buckley, 2014)
      + Medical (Amrein, 2011; del Carmen, 2003)
      + Authorship analyis of 8 million papers - West, 2013
      + Reviewer practice by gender - Grod, 2008
      + recognition for group work - Sarsons, 2017
  
  + Few look at submissions level data to understand source of disparity
      + AGU
      + JAMA (Gilbert & Williams)
      + (Buckley, 2014)
      + eLife
  
  + This study
      + Cummulative data since 2012 to see where ASM stands
      + Build off of previous paper on ASM representation where there indicated a disparity in performance between men & women

#Results

##Papers submitted by women are rejected more frequently than those submitted by men
```{r, plot_rej_rates} 
source("../code/reject_by_gender.R")

auth_type_A #A #cross ASM rejection rates (all genders)
auth_type_B #B #rejection rates by journ (m/w) -- need to highlight instances with diff > 5%
auth_type_C #breakdown of submitted manuscripts (% by gender) according to EJP.decision
auth_type_D #breakdown of C by journal -- need to highlight instances with diff > 5%
```

  + NOTE: to correct for the large discrepency in the participation of women at ASM journals during this time period, all comparisions are made relative to the gender and population in question.
  + direct comparision of the proportion of woman-authored papers that are rejected compared to men, at each author stage
      + middle/first author rejected at similar rates
      + woman-authored (corres/last) rejected more frequently than men
      + there are several journals where the overall trend is repeated and/or amplified (e.g., AAC, IAI, JB, mBio, MCB)
      + greatest effect seen for corresponding authors, use this sub-population to further examine
  + directly compare decisions made on submitted manuscripts by author gender
      + men recieve favorable decisions (accept, revise)
      + most decisions are Accept, Reject, Revise -- look more closely at journal level
      + in most cases (except msphere, msystems, CVI), rejections are more frequently rendered to women than men

##Both men and women editors are more likely to reject papers authored by women than those authored by men
```{r, plot_ed_rej}
source("../code/editorial_rejections.R")#A
source("../code/editorial_rejections_by_journal.R") #E
source("../code/editor_gender_analysis.R")
source("../code/rej_by_inst_type.R")

ed_rejections_A #comparing editorial rejection rates by journal
ed_rejections_E #comparing first decisions by journal
ed_rejections_B #comparing decisions by editor gender
ed_rejections_C #comparing rejection/acceptance rates by instution type
ed_rejections_D #comparing acceptance rates by editor gender, author gender & author institution
```

  + There are two main gatekeeping roles for manuscript decisions: editors & reviewers
  + This section evaluates disparities made by editors during editorial rejections, and decisions following peer review. 
  + Women recieve more editorial rejections than men, proportionally (editorial_rejections.R)
  + may be complications due to geographic bias (eLife paper), restrict to papers submitted in US & examine institution type
  + slight disparity in editorial rejection & acceptance rates according to institution type (rej_by_inst_type.R)
      + greatest difference (4%) occurs for institute/medical schools
      + trend holds for most journals and is >20% diff at JCM (R2s), mBio (Federal), MCB (low research), AEM (industry research) (Supplementary_A)
      + men from institutes/medical schools outperform women >7% for acceptance across all journals 
      + >20% in favor of men at: EC (fed researh), mSphere & JVI (Low research), AEM & MCB (industry research), JCM/JB/JCM (R2 institution) (Supplementary_B)
  + both men & women editors are more likely to reject women (editor_gender_analysis.R)
      + men editors are more likely to make revise only & accept decisions for men
      + women editors highly favor men from medical schools, slightly favor men from R2 & industry
      + men editors favor men from R1, R2 and medical schools, slightly favor women from low & industry research

##Both men and women reviewers more frequently suggest rejection for papers authored by women
```{r, reviewer_rec}
source("../code/rev_score_analysis.R")
reviewer_A
reviewer_B

source("../code/reviewer_gender_analysis.R")
reviewer_C #difference in recommendations by reviewer gender
reviewer_D #difference in acceptance recommendations by reviewer gender & institution type

```

  + rev_score_analysis.R
    + reviewers are more likely to suggest rejections for women as compared to men. No difference in revise decisions (A)
    + men at R1, low research, medical schools & fed research are favored by reviewers over women (B)
  + reviewer_gender_analysis.R
    + both female & male reviewers are more likely to recommend rejection for women. (C)
    + Male reviewers are more likely to accept papers from men (C)
    + women reviewers are more likely to recommend acceptance for women from low research & federal institutions (D)
    + women reviewers are more likely to recommend acceptance for men from R1 & industry instutitions (C)
    + men are more likely to recommend acceptance for women from R2 institutions & favor men from R1 & medical institutes (D)
  
##Factors affecting competency  
```{r, other_factors, echo=FALSE, message=FALSE}
source("../code/time_to_publication.R")
factors_A #days pending at each journal

source("../code/time_at_journ.R")
factors_B #days in peer review system

source("../code/time_to_final_rejection.R") #supplementary D & E

source("../code/impact_analysis.R")
factors_C #cites per month
factors_D #total manuscript reads (HTML+PDF)

source("../code/bias_by_jif.R")
factors_E #no difference in submitted to mbio vs msphere
```
  
  + In addition to rejectance/acceptance rates, other disparate outcomes may occur during the peer review process
  + Time in Peer Reivew (time_to_publication.R)
      + Papers published by women take slightly longer (from submission to ready for publication) than men at some journals (mSphere, mBio, mSystems, CVI, JB, JCM, AEM) despite spending similar amounts of time at ASM journals (A, B)
      + but do not require a greater number of revisions to be accepted than men (Supplementary_C)
      + Papers rejected following review that were submitted by women do not generally take longer (in days) to be rejected (except at mSystems) or have more revisions (Supplementary_D & _E)
      + Decisions are returned to men & women similarly (time_at_journ.R)
  + Peer review is not a linear process, it is cyclical and builds upon its self, may also be disparities in impact (impact_analysis.R)
      + Women tend to receive lower cites per month published than men 
      + HTML/PDF/Abstract views are equivalent 
      + women are more likely to be published in lower JIF journals
      + No disparity in gendered sumissions to high (mBio) vs lower (mSphere) journals (bias_by_jif.R)


  + Notes for plots:
      + highlight instances of >5% difference between genders
      + standardize gender terms to m/w
      + make faceting headers prettier

#Discussion

  + summarize findings
  + fit into context of field -- 
      + few papers have found disparties between rejection rates of men & women
      + first paper (to our knowledge) to collectly examine this issue on 10+ journals using submissions data
      + literature to connect institution type findings
          + compentency? 
          + access to resources
          + "family friendly"/less committed institutions
          + other sterotypes
  + study limitations
      + doesn't include reviewer comments
      + limited journal dataset, difficult to know how other microbiology-publishing journals affect submissions to ASM journals
      + many ASM journals have a narrow focus, & broad scope journals are relatively new -- limits conclusions b/c of journal selection
  + future directions
      + implicit bias training is insufficient
      + consider double-blinded peer review 
          + common practice in social sciences
          + can then be unblinded following decision
      + increase diversity of gatekeepers 
          + increase inclusivity & goodwill/author buy-in
          + women/urm at "fringes" of fields, less likely to be viewed as "competent"

#Data and Methods

###Data

All manuscripts handled by ASM journals (mBio...) that recieved an editorial decision between January 1st, 2012 and August 31st, 2018 were supplied as XML files by ASM's publishing platform, eJP. Data were extracted from the XML documents provided using R statistical software (version 3.4.4) and the `XML` package. Data manipulation was handled using the `tidyverse`, `lubridate`, and `xml2` packages for R. Variables of interest included: the manuscript number assigned to each submission, manuscript type (e.g., full length research, erratum, editorial), category (e.g., microbial ecology), related (previously submitted) manuscripts, versions submitted, dates (e.g., submission, decision), author data (e.g., first, last, and corresponding authorships, total number of authors), reviewer data (e.g., reviewer score, recommendation, editor decision), and person data (names, institutions, country) of the editors, authors, and reviewers. 

It is common practice at ASM journals for manuscripts whose reviewers recommend extensive experimental revisions be given a decision of "reject with resubmission encouraged". If resubmitted, the authors are asked to note the previous (related) manuscript and the resubmission is assigned a new manuscript number. Multiple related manuscripts were tracked together by generating a unique grouped manuscript number based on the recorded related manuscript numbers. This grouped manuscript number served multiple purposes including: tracking a single manuscript through multiple rejections or transfers between ASM journals and to avoid duplicate counts of the same authors. 

###Gender assignment

The gender assignment API genderize.io was used to predict an indivdual's gender based on their given names, and country where possible. The genderize.io platform uses data gathered from social media to predict gender based on given names with the option to include an associated language or country to enhance the odds of succesful prediction. Since all manuscripts are submitted in English, we matched each individuals country against the list of X country names accepted by genderize.io. Using the `GenderGuesser` package for R, all unique given names associated with an accepted country were submitted to the genderize.io API and any names returned without a predictive assignment of either male or female were resubmitted without an associated country. All predictive assignments of either male or female are returned with a probability match of 0.50 or greater. The predicted genders of all given names (with and without an associated country) whose probabilities were greater or equal to our arbitrary success cut off of 0.65 were used to assign predicted gender to the individuals in our dataset. Predicted genders were assigned to individuals in the following order: first names and country, first names, middle names and country, middle names. Individuals who were not assigned a gender are given values of "N/A" and referred to in the analysis as those with an "unknown" gender. The predicted (male/female) gender of editors and senior editors in our dataset was hand validated using Google. 

###Data analysis and presentation


#References

#Supplementary
```{r, Supplementary_1}
source("../code/rej_by_inst_type.R")

Supplementary_A
Supplementary_B

Supplementary_C #number of versions by gender nd journal
Supplementary_D #days from submission to rejection by journal & gender
Supplementary_E #number of versions to rejection by gender & journal
```
