---
title: "Supplementary"
csl: mbio.csl
fontsize: 11pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
always_allow_html: yes
geometry: margin=1.0in
---

```{r supp_setup, include=FALSE}
library(knitr)
library(kableExtra)
library(markdown)
library(rmarkdown)
library(formattable)
library(png)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
```

#Gender representation

\includegraphics{supp_inst_stats.png}

Supplemental Figure . Break down of individuals in each role by their gender and US institute type. Authors are restricted to senior authors (e.g., corresponding, last).

\newpage

#Gender bias
\includegraphics{Supp_inst.png}

Supplemental Figure . Difference in acceptance and rejections by institution type.

\newpage
\includegraphics{Supp_time.png}

Supplemental Figure . Total days and versions between submission and final rejection at each journal.

\newpage
##Gender prediction and assignment

\includegraphics{genderize_method.png}

Supplemental Figure . Schematic of gender prediction and assignment.

\newpage
##Validating gender analysis
```{r genderize_setup, include=FALSE}
source("../code/genderize/b_c_comparison.R")
```

Supplemental Table 1. sensitivity/specificity/accuracy of genderize thresholds. Bolded text denotes the accuracy of the threshold used in all further analyses.
```{r genderize_stats, results='asis', echo=FALSE}  
supp_table_1 <- stats_summary %>% 
  mutate(b_c_data_ascii_pmod85 = ifelse(b_c_data_ascii_pmod85 == "0.9714", 
    cell_spec(b_c_data_ascii_pmod85, bold = T),
    cell_spec(b_c_data_ascii_pmod85, background = "white")),
    b_c_country_data_ascii_pmod85 = ifelse(b_c_country_data_ascii_pmod85 == "0.9695", 
    cell_spec(b_c_country_data_ascii_pmod85, bold = T),
    cell_spec(b_c_country_data_ascii_pmod85, background = "white"))
  ) %>% 
knitr::kable(., format = "latex", digits = 4, col.names = c("Measure", "p0.5", "p0.85", "pmod0.85", "p0.5", "p0.85", "pmod0.85"), escape = F) %>%
  kable_styling() %>% 
  add_header_above(c(" " = 1, "First Names" = 3, "Plus Country Data" = 3)) 

supp_table_1
```

\vspace{40mm}

\includegraphics{impact_equation.png}

Supplemental Figure . Equation for calculating negative bias by genderize. C indicates an individual country. 

\newpage
\includegraphics{Supp_genderize_3.png}

Supplemental Figure . The negative impact of each country on the overall gender prediction of the validation dataset. Number indicates the total number of names associated with each country.

\newpage
\includegraphics{Supp_genderize_4.png}

Supplemental Figure . The negative impact of each country on the overall gender prediction of the full dataset. Number indicates the total number of names associated with each country.

