---
title: "Supplementary"
csl: mbio.csl
fontsize: 11pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
always_allow_html: yes
geometry: margin=1.0in
---

```{r supp_setup, include=FALSE}
library(knitr)
library(kableExtra)
library(markdown)
library(rmarkdown)
library(formattable)
library(png)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
```

#Gender representation

\includegraphics{rep_supp.png}

Supplemental Figure 1. The proportion of (A) potential reviewers at all ASM journals combined, (B) reviewers at each ASM journal, (C) all submitting (dashed line) and publishing (solid line) middle authors, and (D) all submitting and publishing last authors by gender.

\newpage

#Gender bias
\includegraphics{Supp_inst.png}

Supplemental Figure . Difference in acceptance and rejections by institution type.

\newpage
\includegraphics{Supp_time.png}

Supplemental Figure . Total days and versions between submission and final rejection at each journal.

\newpage
##Gender prediction and assignment

\includegraphics{genderize_method.png}

Supplemental Figure . Schematic of gender prediction and assignment.

\newpage
##Validating gender analysis
```{r genderize_setup, include=FALSE}
source("../code/genderize/b_c_comparison.R")
```

Supplemental Table 1. sensitivity/specificity/accuracy of genderize thresholds. Bolded text denotes the accuracy of the threshold used in all further analyses.
```{r genderize_stats, results='asis', echo=FALSE}  
supp_table_1 <- stats_summary %>% 
  mutate(b_c_data_ascii_pmod85 = ifelse(b_c_data_ascii_pmod85 == "0.9714", 
    cell_spec(b_c_data_ascii_pmod85, bold = T),
    cell_spec(b_c_data_ascii_pmod85, background = "white")),
    b_c_country_data_ascii_pmod85 = ifelse(b_c_country_data_ascii_pmod85 == "0.9695", 
    cell_spec(b_c_country_data_ascii_pmod85, bold = T),
    cell_spec(b_c_country_data_ascii_pmod85, background = "white"))
  ) %>% 
knitr::kable(., format = "latex", digits = 4, col.names = c("Measure", "p0.5", "p0.85", "pmod0.85", "p0.5", "p0.85", "pmod0.85"), escape = F) %>%
  kable_styling() %>% 
  add_header_above(c(" " = 1, "First Names" = 3, "Plus Country Data" = 3)) 

supp_table_1
```

\vspace{40mm}

\includegraphics{impact_equation.png}

Supplemental Figure . Equation for calculating negative bias by genderize. C indicates an individual country. 

\newpage
\includegraphics{Supp_genderize_3.png}

Supplemental Figure . The negative impact of each country on the overall gender prediction of the validation dataset. Number indicates the total number of names associated with each country.

\newpage
\includegraphics{Supp_genderize_4.png}

Supplemental Figure . The negative impact of each country on the overall gender prediction of the full dataset. Number indicates the total number of names associated with each country.

