---
title: "Supplementary"
author: "Ada Hagan"
date: "April 17, 2019"
output: html_document
---

```{r supp_setup, include=FALSE}
library(knitr)
library(kableExtra)
library(markdown)
library(rmarkdown)
library(formattable)

source("../code/analysis_functions.R")
source("../code/get_plot_options.R")
source("../code/load_data.R")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

##Gender prediction and assignment
Supplemental Figure 1 - schematic of gender prediction & assignment

##Validating gender analysis
```{r genderize_setup, include=FALSE}
source("../code/genderize/b_c_comparison.R")
```
Supplemental Table 1 - sensitivity/specificity/accuracy of genderize thresholds
```{r genderize_stats, results='asis'}  
stats_summary %>% 
  mutate(b_c_data_ascii_pmod85 = ifelse(b_c_data_ascii_pmod85 == "0.9714", 
    cell_spec(b_c_data_ascii_pmod85, bold = T),
    cell_spec(b_c_data_ascii_pmod85, background = "white")),
    b_c_country_data_ascii_pmod85 = ifelse(b_c_country_data_ascii_pmod85 == "0.9695", 
    cell_spec(b_c_country_data_ascii_pmod85, bold = T),
    cell_spec(b_c_country_data_ascii_pmod85, background = "white"))
  ) %>% 
knitr::kable(., format = "html", table.attr = "style = \"color: black;\"", digits = 4, col.names = c("Measure", "p0.5", "p0.85", "pmod0.85", "p0.5", "p0.85", "pmod0.85"), escape = F) %>%
  kable_styling() %>% 
  add_header_above(c(" " = 1, "First Names" = 3, "Plus Country Data" = 3)) %>% 
  footnote(general = "Bolded text denotes the accuracy of the threshold used in all further analyses")
```
Supplemental Figure 2 - Equation for calculating negative bias by genderize

Supplemental Figure 3 - Validation dataset, country impact
```{r country_impact, message=FALSE}
source("../code/genderize/b_c_country_na.R")

plot_country_impact
```

Supplemental Figure 4 - country impact for full ASM dataset
```{r ASM_genderize, message=FALSE}
source("../code/genderize/ASM_country_na.R")

plot_ASM_country_impact
```

Supplemental Figure 5 
```{r, Supplementary_1}
source("../code/rej_by_inst_type.R")

Supplementary_A
Supplementary_B

Supplementary_C #number of versions by gender nd journal
Supplementary_D #days from submission to rejection by journal & gender
Supplementary_E #number of versions to rejection by gender & journal
```
